Rice yield improvements by plant breeding are offset by decreases over time

### 2017-08-06 output/breed_paper.Rnw
Commit before rearranging the paper

\documentclass[12pt,]{elsarticle}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{longtable,booktabs}
\usepackage{graphicx,grffile, textcomp}
\usepackage{setspace}
\usepackage{etoolbox}
\usepackage{float}
\doublespacing
\usepackage{lineno}
\usepackage{lscape} %% Harvard
\bibliographystyle{model2-names}\biboptions{authoryear}
\pagenumbering{gobble}
%% For figure numbers in supplementary info
%% from http://bytesizebio.net/2013/03/11/adding-supplementary-tables-and-figures-in-latex/
\newcommand{\beginsupplement}{%
  \setcounter{table}{0}
  \renewcommand{\thetable}{S\arabic{table}}%
  \setcounter{figure}{0}
  \renewcommand{\thefigure}{S\arabic{figure}}%
}

@@

Meeting the challenge of feeding an growing population with
increasingly limited resources will require increasing the yield
potential of staple crops, such as rice. Yet many high-yielding,
intensive production systems have experienced slow rates of yield
improvement in recent years despite demonstrated increased yield
potential of new genetics over old in side-by-side comparisons. We
analyzed experimental data from one such system, California (CA) rice
production, in order to better understand what improvements have been
made in the genetic yield potential via plant breeding. Specifically,
the hypothesis that a rice variety's yield erodes over time after the
initial selection was tested as an explanation for how side-by-side
comparisons show increased yield while no substantive increases have
been seen in farmers' fields. We confirmed an apparent increase in
yield in side-by-side comparisons between old and newer varieties when
the yield erosion was not
considered. However, we demonstrate this apparent increase in yield
can be explained by the differences in the time since selection
between cultivars. The impact of yield erosion was estimated to be
29.8 kg ha\tsups{-1} year\tsups{-1} after initial
selection. Once this effect was considered, the apparent yield
advantage of newer cultivars over old was uncertain. This underscores
the importance of breeding to maintain yield and the challenges of
increasing the genetic yield potential.

@@

California (CA) rice production represents one such production
system. Rice yields in CA during the period from 1920 -- 1990
increased rapidly (Fig. \ref{fig:state_yield}), in part due to
improvements in rice genetics (i.e., the ``Green Revolution'') through
the efforts of the California Cooperative Rice Research Foundation
(CCRRF; a collaboration between the University of California, the
USDA-Agricultural Research Service, and farmer-funded
research). However, since 1990 CA state average yields have increased
more slowly (Fig. \ref{fig:state_yield}) despite continuous plant
improvement \citep{ccrrf}. For the last 30 years, each newly developed
cultivar has demonstrated a yield advantage over the previous
cultivars in the Statewide Variety Trials (an annual multi-site
variety trial program that compares experimental entries side-by-side
with ``check'' cultivars). Yet, taken across years, these same reports
show stagnant yields. For example, at the year of release (1985),
cultivar M-202 averaged 11 Mg ha\tsups{-1} \citep[14\%
MC;][]{m202}. In 2015, the reported average yield of the newest released
cultivar M-209 was 10.8 Mg ha\tsups{-1} \citep{m209}. Yet this same
report, M-209 demonstrated a X\% yield advantage over M-202. In other rice
systems, a decrease in a cultivar's yield performance over time has
been observed \citep{Peng1998,
Peng2010, Peng2000, Datta1995}, often termed yield erosion
\citep{Peng1998}. It has been speculated that this is due to the
inability of a selected cultivar's genetics to adapt to changing  biotic and
abiotic conditions. We sought to quantify this phenomenon to better
understand how plant improvement and yield erosion are contributing to
the slow yield improvement in farmers' fields over the last
30 years in CA. Specifically, two competing hypotheses were
tested: (1) the yield advantage of each newly developed cultivars is
greater than previous cultivars, and (2) the erosion of yield over
time is creating the illusion of improved yields (i.e., the new
cultivar is recovering the old benchmark as the older cultivar's yield
erodes).

@@

@
\caption{Average rice yields over the period \Sexpr{min(CA_data$Year)}
  -- \Sexpr{max(CA_data$Year)} for California, USA, as reported by the
  US Dept. of Agriculture National Agricultural Statistics
  Service. Rapid gains in grain yield were experienced from
  approximately 1950 -- 1990, but annual gains since 1990 have been
  slower.}%The line represents the 10 year moving average.}
% Remove 10 year moving average - Bruce does not like. 
\label{fig:state_yield}
\end{figure}

@@

The CCRRF Rice Experiment Station (RES) is located north-west of
Biggs, CA (39.4648, -121.7342), and has been the central location for
CCRRF's efforts to develop improved rice varieties adapted to CA since
1969. The climate at the RES is Mediterranean, characterized by mild
winters during which most of the annual precipitation occurs
(\Sexpr{18*24.5} -- \Sexpr{25*24.5} mm annually) and warm summers
largely free of precipitation events. Soils at the RES are classified
as Esquon-Neerdobe clays with roughly 2\% SOM \citep{ssurgo} in the
top 15cm. As the central location for the CCRRF breeding program, the RES is
researcher-managed (as opposed to farmer-managed trials) and conducts
the bulk of the trials for the Statewide Variety Trials (approximately
3 -- 4 times the number of plots compared to any other location). For
these reasons, we focused our investigation on the RES to avoid
potential complications due to site, climate, and management
differences at the other 6 -- 8 farmer-managed trials per year.

As part of the annual Statewide Variety Trials, newly developed
entries are tested against named and released cultivars (checks) at
the RES in several trials spanning multiple planting dates per
year. For the purposes of this study, we concentrated on released
medium grain cultivars, as medium grain cultivars constitute roughly
90\% of the planted area in CA. These cultivars are typcically planted
in the Statewide Variety Trials for 2 -- 4 years prior to their
official release. For these cultivars planted in
breeding trials at the RES from \Sexpr{min(mod_data[,'year'])} --
\Sexpr{max(mod_data[,'year'])} plot level observations were
collected leading to a final data set of
\Sexpr{format(nrow(mod_data), big.mark = ',')} observations with
\Sexpr{length(unique(mod_data[,'id']))} cultivars represented,
including every publicly released medium grain variety developed by
the CCRRF from \Sexpr{min(mod_data[,'release_yr'])} --
\Sexpr{max(mod_data[,'release_yr'])}. For each cultivar, the official
year of release (typically 3 -- 5 years after introduction to the
trials but prior to widespread commercial seed availability) and the
time elapsed (in years) since release was calculated.

@@

\subsection{Yield erosion}

There is estimated to be an decrease of yield over time once a
cultivar is selected (i.e., yield erosion) (\Sexpr{yrs_coef} kg
ha\tsups{-1} year\tsups{-1}, 90\% CI: \Sexpr{RES_CI[2]} --
\Sexpr{RES_CI[1]}). The oldest cultivar in the trials is estimated to
have lost an average of \Sexpr{round(yrs_coef*-36/1000,1)} Mg
ha\tsups{-1} since release. Comparing two dominate medium grain rice
cultivars M-202 (widely planted from the late 1980s until the early
2000s), and M-206 (the most widely planted CCRRFS cultivar at
approximately 50\% of planted area at the time of writing), the two
cultivars experienced similar trends over time to each other
(Fig. \ref{fig:compare_yrs}) and to the estimate from all released
medium grains (Table \ref{tbl:coef}). At the time of writing, an
estimated \Sexpr{m202_v_206} kg ha\tsups{-1} yield advantage of M-206
over M-202 can be attributed solely to the erosion of M-202's yield
over time (Figs \ref{fig:release} \& \ref{fig:compare_vars}). Further
supporting this estimate, both cultivars yielded similarly at their
respective year of release (Fig. \ref{fig:compare_yrs}), despite the
fact that M-202 was released 19 years prior to M-206
(Fig. \ref{fig:release}, Table \ref{tbl:ids}).

@@

invisible(sapply(1:2, function(i){
    plot(mean ~ year,
         data = plot_data[[i]],
         xlim = c(1985, 2016), ylim = c(7.5,13.5),
         pch = 16, main = gsub("M","M-", tars[i]),
         ylab = "",
         yaxt = "n",
         xlab = "Year",
         cex.axis = 0.8)
    segments(x0 = plot_data[[i]]$year,
             y0 = plot_data[[i]]$mean + plot_data[[i]]$se,
             y1 = plot_data[[i]]$mean - plot_data[[i]]$se)
    if(i == 1){
        axis(2, cex.axis = 0.8,)
        mtext(expression("Yield ( Mg ha"^-1~")"), 2, line = 2.25, cex.lab = 0.6)
    }

    # abline(h = mean(mod_data$yield_kg[(mod_data$id == tars[i] &
    #                                    mod_data$site == "RES")] ),
    #        lty = 1, col = "grey50")
    mod = lm(mean ~ year, data = plot_data[[i]])
    tmp = predict(mod)
    if(i == 1){
        abline(mod)
    }else{
        lines(as.numeric(names(tmp)), tmp)
    }
    abline(h = 11.2, lty = 2)
    text(2000, 7.5, bquote(.(round(coef(mod)[2] * 1000, 1))~kg~ha^-1~year^-1))
}))

@
\caption{The performance of two major cultivars at the CA Rice
  Experiment Station (RES) over the period
  \Sexpr{min(mod_data[,'year'])} -- \Sexpr{max(mod_data[,'year'])}. The
  solid line represents a simple linear regression of yield for each
  cultivar over the period which it was planted at the RES (fit to only
  that cultivar's data), while the dashed line is the published 3-year
  average yield of M-202 at its release (1982-1984). Both cultivars have
  a similar yield at their respective year of release, and similar
  decreases over time. The slope for the individual regressions are
  comparable to the slope estimated from all medium grain cultivars
  (\Sexpr{yrs_coef} kg ha\tsups{-1} year\tsups{-1}).}
\label{fig:compare_yrs}
\end{figure}

@@

The full model which corrected for both the year of release and yield
erosion was estimated to have better out-of-sample predictive accuracy
compared to reduced models containing only one predictor. As estimated
by the best model, there is no conclusive evidence that the yields of
medium grain rice cultivars selected and released by CCRRF's breeding
program have, on average, increased over the last 30 years. The
estimated change in grain yield per year in the CCRRF's breeding
trials at the RES was estimated to be a negligible
\Sexpr{release_yr_coef} kg ha\tsups{-1} year\tsups{-1}, with both
small positive and negative trends within the credible interval (90\%
CI: \Sexpr{rel_CI[1]} -- \Sexpr{rel_CI[2]})
(Fig. \ref{fig:release}b.). Hence, the evidence here is consistent
with little to no increase in yield potential over the study
period. When yield erosion is not corrected for in the comparison
(i.e., omitted from the model), there is an apparent small increase
per year (approximately \Sexpr{only_release_coef} kg ha\tsups{-1}
year\tsups{-1}; 90\% CI \Sexpr{only_release_CI[1]} --
\Sexpr{only_release_CI[2]}) (Fig. \ref{fig:release}a.), though this
estimate ignores the confounding effect of yield erosion over time and
performed worse in out-of-sample predictive accuracy.

After correcting for the other variables, some cultivars are estimated
to have greater yield compared to the overall average. The marginal
estimate for the performance of a cultivar averaged over the
uncertainty in the other parameters in the model (i.e., the random
effect) shows three cultivars with higher than average yields:
\Sexpr{paste(gsub('M','M-',rev(row.names(id_est))[c(1,2,4)]),collapse=',')}
(Figs. \ref{fig:release} \& \ref{fig:ids_comp}). A fourth, M-209,
might also have higher than average yield, though the CI for M-209
overlaps with zero (Fig. \ref{fig:ids_comp}).

<<fig=TRUE, echo=FALSE, width = 9, height = 6>>=
yrs <- 1975:2025

@@

par(mfrow = c(1,2),
    mar = c(4,0,2,0.75),
    oma = c(4,4,0.5,0.5))

@@

@@

@
% Change from a./b. to accounting for erosion, More details about how
% yield was corrected - 
\caption{The relationship between yield and the release year of medium
  grain rice cultivars developed by the California Cooperative Rice
  Research Foundation (California, USA), for the last 6 years (\Sexpr{min(comp_yrs)} -- \Sexpr{max(comp_yrs)}) as estimated by two methods;
  without correcting for yield erosion over time (left) and 
  correcting for yield erosion over time (right). The trend without correcting
  for yield erosion (left) was estimated by omitting the year of each
  cultivar's release from the model. Yield erosion was corrected for by
  including the years since release (right) in the model.}
\label{fig:release}
\end{figure}

@@

We conclude that yield erosion has decreased the yield of cultivars
released by the CCRRF over the last 30+ years. Biotic and abiotic
forces have been speculated to be responsible for this
yield erosion since the cultivar's genetics remain
``frozen'' at the time of selection but the conditions the cultivar
may experience over time are continuously changing \citep{Peng1998,
Peng2000, Peng2010}. The trend observed here is consistent with
previous reports of the erosion of genetic gains of cultivar IR8 at
IRRI \citep{Peng1998, Peng2000}, suggesting the
erosion of yield over time is a phenomenon that widely affects rice
breeding programs. This highlights both the need for maintenance
breeding and the difficulty of increasing genetic yield potential,
especially in systems such as CA where yield is approaching the
physiological limit \citep{espe16a}. Without maintenance breeding
continuously selecting cultivars well adapted to current conditions,
CA rice production systems are predicted to have lost an estimated
\Sexpr{round(yrs_coef * (-36/1000), 1)} Mg ha\tsups{-1} grain yield
over the study period.

Overall, we do not find conclusive evidence of increased genetic yield
potential in medium grain rice cultivars released by CCRRF's breeding
improvement program over the period \Sexpr{min(mod_data[,'year'])} to
\Sexpr{max(mod_data[,'year'])}. After correcting yield comparisons for
yield erosion over time, the estimated yield improvement spans from
\Sexpr{rel_CI[1]} -- \Sexpr{rel_CI[2]} kg ha\tsups{-1} year\tsups{-1}.
Even when the comparison is not corrected for yield erosion, the
estimated improvements over time (\Sexpr{only_release_coef} kg
ha\tsups{-1} year\tsups{-1}; Fig. \ref{fig:release}a.) are still much
less than previous estimates of 37 -- 60 kg ha\tsups{-1}
year\tsups{-1} \citep{Omar2014}. We conclude that the breeding program
has likely been maintaining but not substantially improving yield
potential in the face of changing biotic and abiotic pressures.

However, the goal of a breeding program is to not only maintain the
benchmark yield of the previous ``standard'' but to surpass it. Given
the estimated yield decreases over time, to hit an ambitious target of
0.5 Mg ha\tsups{-1} (approximately 5\%) increase in yield per release
(with roughly 4 -- 5 years between releases), a new cultivar needs to
yield an additional \Sexpr{round(yrs_coef*-4,-1)} --
\Sexpr{round(yrs_coef*-5,-1)} kg ha\tsups{-1} in side-by-side
comparison on top of the 0.5 Mg ha\tsups{-1} target improvement. Often
realized side-by-side yield improvements are closer to 2 -- 3\% per
release, in which case no absolute improvements are made beyond
maintenance. The evidence here is consistent with the latter situation
occurring in CA, further suggesting that the genetic yield potential
has not increased substantially in the last 30+ years
(Fig. \ref{fig:release}).

@@

@ 
\caption{A conceptual diagram highlighting three scenarios:
  (1) maintenance breeding, where losses over time are offset by
  improvements via plant improvement, (2) no maintenance breeding,
  where grain yield slowly erodes over time, and (3) absolute
  improvement in grain yield, where gains in grain yield exceed the
  losses over time.}
\label{fig:theory}
\end{figure}

California state average yields have increased slowly
even as little to no gains have been seen in the breeding trials
(Fig. \ref{fig:state_yield}). The evidence here suggests mechanisms
other than increased genetic yield potential are responsible for these gains
(Fig. \ref{fig:compare_yrs}). Improvement in agronomic management,
including more efficient nutrient management \citep{Lundy2012,
Lundy2015, linquist2009}, weed management \citep{Saito2008,
Pittelkow2012, linquist2008}, etc. may all be contributing to
increases in the state average yield. However, these gains are less
than those experienced during the ``Green revolution'', and by their
nature agronomic gains cannot surpass the genetic yield potential
\citep{Fischer2015}. Judging from yields achieved at the RES
(Figs. \ref{fig:RES_year} \& \ref{fig:compare_yrs}), a
researcher-managed production system implementing best management
practices in more controlled situations compared to farmers' fields, state
average yields (approximately 10 Mg ha\tsups{-1} at the time of
writing; Fig. \ref{fig:state_yield}) may be approaching the current
genetic limit. Assuming that the level of management possible in small
experimental plots cannot be attained in farmers' fields and hence
experimental yields represent the attainable yield ceiling
\citep{Fischer2015}, agronomic gains in farmer's fields are nearing
their potential, consistent with physiological model-based estimates
of the attainable yield gap in CA rice production systems
\citep{espe16a}.

>>>>

\documentclass[12pt,]{elsarticle}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{longtable,booktabs}
\usepackage{comment}
\usepackage{graphicx,grffile, textcomp}
\usepackage{setspace}
\usepackage{etoolbox}
\usepackage{float}
\doublespacing
\usepackage{lineno}
\usepackage{lscape} %% Harvard
\bibliographystyle{model2-names}\biboptions{authoryear}
\pagenumbering{gobble}
%% For figure numbers in supplementary info
%% from http://bytesizebio.net/2013/03/11/adding-supplementary-tables-and-figures-in-latex/
\newcommand{\beginsupplement}{%
  \setcounter{table}{0}
  \renewcommand{\thetable}{S\arabic{table}}%
  \setcounter{figure}{0}
  \renewcommand{\thefigure}{S\arabic{figure}}%
}

@@

Meeting the challenge of feeding an growing population with
increasingly limited resources will require increasing the yield
potential of staple crops, such as rice. Yet many high-yielding,
intensive production systems have experienced slow rates of yield
improvement in recent years despite demonstrated increased yield
potential of new genetics over old in side-by-side comparisons. We
analyzed experimental data from one such system, California (CA) rice
production, in order to better understand what improvements have been
made in the genetic yield potential via plant breeding. Specifically,
the hypothesis that a rice variety's yield erodes over time after the
initial selection was tested as an explanation for how side-by-side
comparisons show increased yield while no substantive increases have
been seen in farmers' fields. Side-by-side comparisons under yield
erosions will show ``false'' increases due to the older cultivar
decreasing, not due to increased genetic yield potential by the newer
cultivar. We confirmed an apparent increase in yield in side-by-side
comparisons between old and newer varieties when the yield erosion was
not considered. However, we demonstrate this apparent increase in
yield can be explained by the differences in the time since selection
between cultivars. The impact of yield erosion was estimated to be
29.3 kg ha\tsups{-1} year\tsups{-1} (90\% credible interal -4.4 to
-53.3) after initial selection. Once this effect was considered, the
apparent yield advantage of newer cultivars over old was uncertain
(-3.3 kg ha\tsups{-1} year\tsups{-1}, 90\% credible interval -36.1 to
31.5). This underscores the importance of breeding to maintain yield
and the challenges of increasing the genetic yield potential.

@@

California (CA) rice production represents one such production
system. Rice yields in CA during the period from 1920 -- 1990
increased rapidly (Fig. \ref{fig:state_yield}), in part due to
improvements in rice genetics (i.e., the ``Green Revolution'') through
the efforts of the California Cooperative Rice Research Foundation
(CCRRF; a collaboration between the University of California, the
USDA-Agricultural Research Service, and farmer-funded
research). However, since 1990 CA state average yields have increased
more slowly (Fig. \ref{fig:state_yield}) despite continuous plant
improvement \citep{ccrrf}. For the last 30 years, each newly developed
cultivar has demonstrated a yield advantage over the previous
cultivars in the Statewide Variety Trials (an annual multi-site
variety trial program that compares experimental entries side-by-side
with ``check'' cultivars). Yet, taken across years, these same reports
show stagnant yields. For example, at the year of release (1985),
cultivar M-202 averaged 11 Mg ha\tsups{-1} \citep[14\%
MC;][]{m202}. In 2015, the reported average yield of the newest released
cultivar M-209 was 10.8 Mg ha\tsups{-1} \citep{m209}. Yet this same
report, M-209 demonstrated a X\% yield advantage over M-202. In other rice
systems, a decrease in a cultivar's yield performance over time has
been observed \citep{Peng1998,
Peng2010, Peng2000, Datta1995}, often termed yield erosion
\citep{Peng1998}. It has been speculated that this is due to the
inability of a selected cultivar's genetics to adapt to changing  biotic and
abiotic conditions. We sought to quantify this phenomenon to better
understand how plant improvement and yield erosion are contributing to
the slow yield improvement in farmers' fields over the last
30 years in CA. Specifically, two competing hypotheses were
tested: (1) the genetic yield advantage has increased over the last
30+ years, and (2) the genetic yield advantage has not increased over
the last 30+ and the erosion of yield over
time is creating the illusion of improved yields (i.e., the new
cultivar is recovering the old benchmark as the older cultivar's yield
erodes).

@@

@
\caption{Average rice yields over the period \Sexpr{min(CA_data$Year)}
  to \Sexpr{max(CA_data$Year)} for California, USA, as reported by the
  US Dept. of Agriculture National Agricultural Statistics
  Service. Rapid gains in grain yield were experienced from
  approximately 1950 -- 1990, but annual gains since 1990 have been
  slower.}%The line represents the 10 year moving average.}
% Remove 10 year moving average - Bruce does not like. 
\label{fig:state_yield}
\end{figure}

@@

The CCRRF Rice Experiment Station (RES) is located north-west of
Biggs, CA (39.4648, -121.7342), and has been the central location for
CCRRF's efforts to develop improved rice varieties adapted to CA since
1969. The climate at the RES is Mediterranean, characterized by mild
winters during which most of the annual precipitation occurs
(\Sexpr{18*24.5} to \Sexpr{25*24.5} mm annually) and warm summers
largely free of precipitation events. Soils at the RES are classified
as Esquon-Neerdobe clays with roughly 2\% SOM \citep{ssurgo} in the
top 15cm. As the central location for the CCRRF breeding program, the RES is
researcher-managed (as opposed to farmer-managed trials) and conducts
the bulk of the trials for the Statewide Variety Trials (approximately
3 to 4 times the number of plots compared to any other location). For
these reasons, we focused our investigation on the RES to avoid
potential complications due to site, climate, and management
differences at the other 6 to 8 farmer-managed trials per year.

As part of the annual Statewide Variety Trials, newly developed
entries are tested against named and released cultivars (checks) at
the RES in several trials spanning multiple planting dates per
year. For the purposes of this study, we concentrated on released
medium grain cultivars, as medium grain cultivars constitute roughly
90\% of the planted area in CA. These cultivars are typcically planted
in the Statewide Variety Trials for 2 to 4 years prior to their
official release. For these cultivars planted in
breeding trials at the RES from \Sexpr{min(mod_data[,'year'])} --
\Sexpr{max(mod_data[,'year'])} plot level observations were
collected leading to a final data set of
\Sexpr{format(nrow(mod_data), big.mark = ',')} observations with
\Sexpr{length(unique(mod_data[,'id']))} cultivars represented,
including every publicly released medium grain variety developed by
the CCRRF from \Sexpr{min(mod_data[,'release_yr'])} --
\Sexpr{max(mod_data[,'release_yr'])}. For each cultivar, the official
year of release (typically 3 -- 5 years after introduction to the
trials but prior to widespread commercial seed availability) and the
time elapsed (in years) since release was calculated.

@@

@@

invisible(sapply(1:2, function(i){
    plot(mean ~ year,
         data = plot_data[[i]],
         xlim = c(1985, 2016), ylim = c(7.5,13.5),
         pch = 16, main = gsub("M","M-", tars[i]),
         ylab = "",
         yaxt = "n",
         xlab = "Year",
         cex.axis = 0.8)
    segments(x0 = plot_data[[i]]$year,
             y0 = plot_data[[i]]$mean + plot_data[[i]]$se,
             y1 = plot_data[[i]]$mean - plot_data[[i]]$se)
    if(i == 1){
        axis(2, cex.axis = 0.8,)
        mtext(expression("Yield ( Mg ha"^-1~")"), 2, line = 2.25, cex.lab = 0.6)
    }
    # abline(a = ((fixef(RES_full)[1] * scl) + center) / 1000 -
    #            (2000 * yrs_coef/1000),
    #        b = yrs_coef/1000,
    #        lty = 1)
    lines(as.numeric(rownames(plot_data[[i]])),
          11.2 + (1:nrow(plot_data[[i]])) * (yrs_coef/1000))
    # abline(a = ((fixef(RES_full)[1] * scl) + center) / 1000,
           # b = yrs_coef/1000)

    # abline(h = mean(mod_data$yield_kg[(mod_data$id == tars[i] &
    #                                    mod_data$site == "RES")] ),
    #        lty = 1, col = "grey50")
    # mod = lm(mean ~ year, data = plot_data[[i]])
    # tmp = predict(mod)
    # if(i == 1){
    #     abline(mod)
    # }else{
    #     lines(as.numeric(names(tmp)), tmp)
    # }
    abline(h = 11.2, lty = 2)
    # text(2000, 7.5, bquote(.(round(coef(mod)[2] * 1000, 1))~kg~ha^-1~year^-1))
}))

@
\caption{The performance of two major cultivars at the CA Rice
Experiment Station (RES) over the period
\Sexpr{min(mod_data[,'year'])} -- \Sexpr{max(mod_data[,'year'])}. The
solid line is the estimated yield erosion (\Sexpr{yrs_coef} kg
ha\tsups{-1} year\tsups{-1}), while the dashed line is the published
3-year average yield of M-202 at its release (1982-1984). Both
cultivars have a similar yield at their respective year of release,
and similar decreases over time.}
\label{fig:compare_yrs}
\end{figure}

@@

The full model which corrected for both the year of release and yield
erosion was estimated to have better out-of-sample predictive accuracy
compared to reduced models containing only one predictor. As estimated
by the best model, there is no conclusive evidence that the yields of
medium grain rice cultivars selected and released by CCRRF's breeding
program have, on average, increased over the last 30 years. The
estimated change in grain yield per year in the CCRRF's breeding
trials at the RES was estimated to be a negligible
\Sexpr{release_yr_coef} kg ha\tsups{-1} year\tsups{-1}, with both
small positive and negative trends within the credible interval (90\%
CI: \Sexpr{rel_CI[1]} to \Sexpr{rel_CI[2]})
(Fig. \ref{fig:w_erosion}). Hence, the evidence here is consistent
with little to no increase in yield potential over the study
period. When yield erosion is not corrected for in the comparison
(i.e., omitted from the model), there is an apparent small increase
per year (approximately \Sexpr{only_release_coef} kg ha\tsups{-1}
year\tsups{-1}; 90\% CI \Sexpr{only_release_CI[1]} to
\Sexpr{only_release_CI[2]}) (Fig. \ref{fig:wo_erosion}), though this
estimate ignores the confounding effect of yield erosion over time and
performed worse in out-of-sample predictive accuracy.

After correcting for the other variables, some cultivars are estimated
to have greater yield compared to the overall average. The marginal
estimate for the performance of a cultivar averaged over the
uncertainty in the other parameters in the model (i.e., the random
effect) shows three cultivars with higher than average yields:
\Sexpr{paste(gsub('M','M-',rev(row.names(id_est))[c(1,2,4)]),collapse=',')}
(Figs. \ref{fig:w_erosion} \& \ref{fig:ids_comp}). A fourth, M-209,
might also have higher than average yield, though the CI for M-209
overlaps with zero (Fig. \ref{fig:ids_comp}).

\begin{comment}
\begin{figure}[H]

<<fig=TRUE, echo=FALSE, width = 6, height = 5>>=
yrs <- 1975:2025

@@

# par(mfrow = c(1,2),
#     mar = c(4,0,2,0.75),
#     oma = c(4,4,0.5,0.5))

@@

@
\caption{The relationship between yield and the release year of medium
grain rice cultivars developed by the California Cooperative Rice
Research Foundation (California, USA), for the last 6 years
(\Sexpr{min(comp_yrs)} -- \Sexpr{max(comp_yrs)}) as estimated without
correcting for yield erosion over time.}
\label{fig:wo_erosion}
\end{figure}
\end{comment}

\begin{figure}[H]
<<fig=TRUE, echo=FALSE, width = 6, height = 5>>=

@@

@
% Change from a./b. to accounting for erosion, More details about how
% yield was corrected - 
\caption{The relationship between yield and the release year of medium
grain rice cultivars developed by the California Cooperative Rice
Research Foundation (California, USA), for the last 6 years
(\Sexpr{min(comp_yrs)} -- \Sexpr{max(comp_yrs)}) as estimated by 
correcting for yield erosion over time. Yield erosion was corrected
for by adjusting each observed yield by estimated amount lost 
  (\Sexpr{yrs_coef} kg ha\tsups{-1} year\tsups{-1} after release).}
\label{fig:w_erosion}
\end{figure}

\subsection{Yield erosion}
\begin{figure}[H]

There is estimated to be an decrease of yield over time once a
cultivar is selected (i.e., yield erosion) (\Sexpr{yrs_coef} kg
ha\tsups{-1} year\tsups{-1}, 90\% CI: \Sexpr{RES_CI[2]} to
\Sexpr{RES_CI[1]}). The oldest cultivar in the trials is estimated to
have lost an average of \Sexpr{round(yrs_coef*-36/1000,1)} Mg
ha\tsups{-1} since release. Comparing two dominate medium grain rice
cultivars M-202 (widely planted from the late 1980s until the early
2000s), and M-206 (the most widely planted CCRRFS cultivar at
approximately 50\% of planted area at the time of writing), the two
cultivars experienced similar trends over time to each other
(Fig. \ref{fig:compare_yrs}) and to the estimate from all released
medium grains (Table \ref{tbl:coef}). At the time of writing, an
estimated \Sexpr{m202_v_206} kg ha\tsups{-1} yield advantage of M-206
over M-202 can be attributed solely to the erosion of M-202's yield
over time (Figs \ref{fig:w_erosion} \& \ref{fig:compare_yrs}). Further
supporting this estimate, both cultivars yielded similarly at their
respective year of release (Fig. \ref{fig:compare_yrs}), despite the
fact that M-202 was released 19 years prior to M-206
(Fig. \ref{fig:compare_yrs}, Table \ref{tbl:ids}).

@@

We estimate that yield erosion has decreased the yield of medium
grain rice cultivars released by the CCRRF over the last 30+ years by
an average of \Sexpr{yrs_coef} Mg ha\tsups{-1},
potentially offsetting any gains made by the breeding program. Biotic
and abiotic forces have previously been speculated to be responsible
for yield erosion, since the cultivar's genetics remain stable from the
time of selection but the conditions the cultivar may experience over
time are continuously changing \citep{Peng1998, Peng2000,
Peng2010}. The trend observed here is consistent with previous reports
of the erosion of genetic gains of cultivar IR8 at IRRI
\citep{Peng1998, Peng2000}, suggesting the erosion of yield over time
is a phenomenon that widely affects rice breeding programs. This
highlights both the need for maintenance breeding and the difficulty
of increasing genetic yield potential, especially in systems such as
CA where yield is approaching the physiological limit
\citep{espe16a}. Without maintenance breeding continuously selecting
cultivars well adapted to current conditions, CA rice production
systems are predicted to have lost an estimated
\Sexpr{round(yrs_coef*(-36/1000),1)} Mg ha\tsups{-1} grain yield over
the study period.

Overall, we do not find conclusive evidence of increased genetic yield
potential in medium grain rice cultivars released by CCRRF's breeding
improvement program over the period \Sexpr{min(mod_data[,'year'])} to
\Sexpr{max(mod_data[,'year'])}. After correcting yield comparisons for
yield erosion over time, the estimated yield improvement spans from
\Sexpr{rel_CI[1]} to \Sexpr{rel_CI[2]} kg ha\tsups{-1}
year\tsups{-1}. This suggests that both increases and decreases in the
genetic yield potential over the study period. However, even when the
comparison is not corrected for yield erosion, the estimated
improvements over time (\Sexpr{only_release_coef} kg ha\tsups{-1}
year\tsups{-1}; Fig. \ref{fig:wo_erosion}) are still much less than
previous estimates of 37 to 60 kg ha\tsups{-1} year\tsups{-1}
\citep{Omar2014}. We conclude that the breeding program has likely
been maintaining but not substantially improving yield potential in
the face of changing biotic and abiotic pressures.

However, a goal of many breeding programs is to not only maintain the
benchmark yield of the previous ``standard'' but to surpass it,
especially if we are to meet the target of 50\% increased food
production by 2050 \citep{Foley2011a}. Given the estimated yield
erosion, to hit an ambitious target of 0.5 Mg ha\tsups{-1}
(approximately 5\%) increase in yield per release (with roughly 4 to 5
years between releases), a new cultivar needs to yield an additional
\Sexpr{round(yrs_coef*-4,-1)} to \Sexpr{round(yrs_coef*-5,-1)} kg
ha\tsups{-1} in side-by-side comparison on top of the 0.5 Mg
ha\tsups{-1} target improvement. Often realized side-by-side yield
improvements are closer to 2 to 3\% per release, in which case no
absolute improvements are made beyond maintenance. The evidence here
is consistent with the latter situation occurring in CA, further
suggesting that the genetic yield potential has not increased
substantially in the last 30+ years (Fig. \ref{fig:w_erosion}).

@@

@ 
\caption{A conceptual diagram highlighting three scenarios: (1)
maintenance breeding, where losses over time are offset by
improvements via plant improvement, (2) no maintenance breeding, where
grain yield slowly erodes over time (as estimated from this study),
and (3) absolute improvement in grain yield (based on ``typical''
yield goals of 3\% increase per release), where gains in grain yield
exceed the losses over time.}
\label{fig:theory}
\end{figure}

California state average yields have increased slowly
even as little to no gains have been seen in the breeding trials
(Fig. \ref{fig:state_yield}). The evidence here suggests mechanisms
other than increased genetic yield potential are responsible for these gains
(Fig. \ref{fig:compare_yrs}). Improvement in agronomic management,
including more efficient nutrient management \citep{Lundy2012,
Lundy2015, linquist2009}, weed management \citep{Saito2008,
Pittelkow2012, linquist2008}, etc. may all be contributing to
increases in the state average yield. However, these gains are far less
than those experienced during the preceeding era, and by their
nature agronomic gains cannot surpass the genetic yield potential
\citep{Fischer2015}. Judging from yields achieved at the RES
(Figs. \ref{fig:RES_year} \& \ref{fig:compare_yrs}), a
researcher-managed production system implementing best management
practices in more controlled situations compared to farmers' fields, state
average yields (approximately 10 Mg ha\tsups{-1} at the time of
writing; Fig. \ref{fig:state_yield}) may be approaching the current
genetic limit. Assuming that the level of management possible in small
experimental plots cannot be attained in farmers' fields and hence
experimental yields represent the attainable yield ceiling
\citep{Fischer2015}, agronomic gains in farmer's fields are nearing
their potential, consistent with physiological model-based estimates
of the attainable yield gap in CA rice production systems
\citep{espe16a}.

There are several cultivars that were estimated to have better than
average yields, even after correcting for yield erosion. However,
these cultivars (M-201, M-204, M-205, and possibly M-209;
Fig. \ref{fig:ids_comp})) may not be widely adapted. It is known that
cool temperatures during the booting stage can have a large impact on
grain yields \citep{espe2017}, and it is also known that the RES
experiences warmer temperatures than many other rice growing areas in
CA \citep{espe2017}. These cultivars may perform well at the RES but
not widely in CA. Therefore, it should be noted that while our
investigation has been concerned about increasing yield potential, the
estimates here might not capture yield stability. Yield stability is
notoriously difficult to quantify \citep[e.g.,][]{lobell2011}, hence
further investigation is needed to explore how yield stability has
changed over the study period.
